systemfit: A Package for Estimating\\ Systems of Simultaneous Equations in \proglang{R}

### 2007-12-06 paper/trunk/systemfit_other.tex
revised (sub)section "Testing linear restrictions": added derivation of F test and Wald tests, the description of these tests is more clear, more elaborate and fully correct now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@884 591db2bd-e0a0-4d22-8c19-2192dbb7da31

Linear restrictions can be tested by an $F$ test, Wald test or
likelihood ratio (LR) test.

The $F$ statistic for systems of equations is
\begin{equation}
F = \frac{
   ( R \bHat - q )'
   ( R ( X' ( \SHat \otimes I )^{-1} X )^{-1} R' )^{-1}
   ( R \bHat - q ) /
   j
}{
   \uHat' ( \SHat \otimes I )^{-1} \uHat /
   ( G \cdot T - K )
} \; ,
\end{equation}
where $j$ is the number of restrictions,
$G$ is the number of equations,
$T$ is the number of observations per equation,
$K$ is the total number of estimated coefficients, and
$\SHat$ is the (estimated) residual covariance matrix
used in the estimation.
Under the null hypothesis, $F$ has an approximate $F$ distribution
with $j$ and $G \cdot T - K$ degrees of freedom
\citep[p.~314]{theil71}.

The Wald statistic for systems of equations is
\begin{equation}
W =
   ( R \bHat - q )'
   ( R \, \COVHat [ \bHat ] \, R' )^{-1}
   ( R \bHat - q ) .
\end{equation}
Asymptotically, $W$ has a $\chi^2$
distribution with $j$ degrees of freedom
under the null hypothesis
\citep[p.\ 347]{greene03}.

>>>>

Linear restrictions can be tested by an $F$ test, two Wald tests and
a likelihood ratio (LR) test.

The $F$ statistic for systems of equations is
\begin{equation}
F = \frac{
   ( R \bHat - q )'
   ( R ( X' ( \Sigma \otimes I )^{-1} X )^{-1} R' )^{-1}
   ( R \bHat - q ) /
   j
}{
   \uHat' ( \Sigma \otimes I )^{-1} \uHat /
   ( G \cdot T - K )
} \; ,
\label{eq:f-test}
\end{equation}
where $j$ is the number of restrictions,
$K$ is the total number of estimated coefficients,
and all other variables are as defined before
\citep[p.~314]{theil71}.
Under the null hypothesis, $F$ is $F$ distributed
with $j$ and $G \cdot T - K$ degrees of freedom.

However, $F$ in equation~(\ref{eq:f-test}) cannot be computed,
because $\Sigma$ is generally unknown.
As a solution, \citet[p.~314]{theil71} proposes
to replace the unknown $\Sigma$ in~(\ref{eq:f-test})
by the estimated covariance matrix $\SHat$.
\begin{equation}
\hat{F} = \frac{
   ( R \bHat - q )'
   ( R ( X' ( \SHat \otimes I )^{-1} X )^{-1} R' )^{-1}
   ( R \bHat - q ) /
   j
}{
   \uHat' ( \SHat \otimes I )^{-1} \uHat /
   ( G \cdot T - K )
}
\label{eq:f-test-theil}
\end{equation}
Asymptotically, $\hat{F}$ has the same distribution as $F$
in equation~(\ref{eq:f-test}),
because the numerator of~(\ref{eq:f-test-theil})
converges in probability to the numerator of~(\ref{eq:f-test})
and the denominator of~(\ref{eq:f-test-theil})
converges in probability to the denominator of~(\ref{eq:f-test})
\citep[p.~402]{theil71}.

Furthermore, the denominators of both~(\ref{eq:f-test})
and~(\ref{eq:f-test-theil}) converge in probability to~$1$.
Taking this into account and applying equation~(\ref{eq:cov-ols-wls-sur}),
we obtain the usual $F$~statistic of the Wald test.
\begin{equation}
\hat{\hat{F}} = \frac{
   ( R \bHat - q )'
   ( R \, \COVHat \left[ \bHat \right] R' )^{-1}
   ( R \bHat - q )
}{
   j
}
\label{eq:f-test-wald}
\end{equation}
Under the null hypotheses, also $\hat{\hat{F}}$ is
asymptotically $F$ distributed
with $j$ and $G \cdot T - K$ degrees of freedom.

Multiplying equation~(\ref{eq:f-test-wald}) with $j$,
we obtain the usual $\chi^2$ statistic for the Wald test
\begin{equation}
W =
   ( R \bHat - q )'
   ( R \, \COVHat [ \bHat ] \, R' )^{-1}
   ( R \bHat - q ) .
\end{equation}
Asymptotically, $W$ has a $\chi^2$
distribution with $j$ degrees of freedom
under the null hypothesis
\citep[p.\ 347]{greene03}.
